Meta-analytic approaches and effect sizes to account for 'nuisance heterogeneity' in comparative physiology

### 2021-09-17 MS/ms.Rmd
Fixed a few more thigs

Meta-analysis has emerged as the 'gold-standard' across various disciplines for quantitative synthesis and is becoming increasingly prominent in the field of comparative physiology  [@Gurevitch2018; @Nakagawa2017; @Glass2015]. Meta-analyses answer research questions by synthesising research results and identifying sources of variation across studies [@Nakagawa2017; @Gurevitch1999; @Gurevitch2018; @Cooper2009; @Arnqvist1995; @Koricheva2013; @Borenstein2019]. Ideally, meta-analyses are part of a systematic review and methods are carefully reported upon to ensure the review and data collation are transparent and reproducible [i.e., see the Preferred Reporting Items for Systematic Reviews and Meta-Analyses in Ecology and Evolutionary Biology (PRISMA-EcoEvo) checklist, @ODea2021].

A meta-analysis can have three goals: 1) to provide an overall mean estimate of a treatment effect or relationship, 2) to quantify effect size variance and understand key drivers explaining differences in effects across studies and 3) attempt to identify research gaps and publication biases [@Nakagawa2017; @Gurevitch1999; @Gurevitch2018; @Cooper2009; @Koricheva2013; @Borenstein2019]. Meta-analysing independent studies provides greater statistical power and precision than what any individual study on its own would be able to provide -- particularly given that most empirical studies are already under-powered in many areas of biology [@Forstmeier2017; @Jennions2003; @Button2013]. By expressing study effects on a common scale (i.e., standardised effect size) we can gain broader insight into the direction and efficacy of a particular treatment effect or the strength of a relationship between two variables of interest [@Gurevitch2018; @Koricheva2013]. Meta-analyses have already provided comparative physiologists with powerful insights on pressing global challenges -- from testing whether physiological plasticity can buffer organisms against climate change [e.g, @Seebacher2015] to the degree to which endocrine disrupting chemicals, such as bis-phenol A (BPA), impact aquatic organisms [e.g., @WuSeebacher2020].

@@

We included comparative studies / meta-analyses that were focused on physiological questions. Narrative reviews and comments were not considered because they did not collate quantitative data from the literature. Studies needed to have collected physiological trait data, such as hormone levels, thermal physiological parameters (CTmax, CTmin, thermal preferences, thermal breadth, LT50, lethal thresholds), metabolic rates (routine, maximum, and minimum metabolic rate, aerobic scope), enzyme reaction rates and concentrations, oxidative stress measures and consequences (e.g., DNA damage, teleomere attrition), performance measurements (e.g., heart rates, sprint speed, swim speed, growth rates), body composition (e.g., fatty acid profiles) and immune traits (e.g., cell function, antibodies). Studies were also included if they measured the relationship between physiological traits and body size (e.g., metabolic scaling), reproductive allocation, or survival across taxa. Studies on behavior were only included if the synthesis was explicitly interested in the proximate physiological drivers of behavioral variation. We also included meta-analyses that quantified plasticity in physiological traits in response to environmental variation. While many studies in comparative physiology collect data as part of a comparative analysis, we focus here only on studies collating data from existing literature because this is the major data source in quantitative syntheses. We excluded quantitative syntheses interested in phenology, bio-mechanics, sexual size dimorphism (i.e., except if the study was interested in sexual differences in physiological traits), those focused on validating physiological methodology, or characterising population genetic structure, gene transcription, and quantitative genetics. There were also meta-analyses that focused on colour evolution which varied in their goals but were most often not focused on physiological questions. As such, these studies were also excluded. We restricted our focus to animal studies because most plant synthesis papers were focused on ecological (e.g., quantified changes in abundance and distribution of species) or agricultural questions (e.g., studies on soil microbial function, nitrogen and phosphorus capture, methane output, and greenhouse gas emissions).

>>>>

Meta-analysis has emerged as the 'gold-standard' across various disciplines for quantitative synthesis and is becoming increasingly prominent in the field of comparative physiology  [@Gurevitch2018; @Nakagawa2017; @Glass2015]. Meta-analyses answer research questions by synthesising research results and identifying sources of variation across studies [@Nakagawa2017; @Gurevitch1999; @Gurevitch2018; @Cooper2009; @Arnqvist1995; @Koricheva2013; @Borenstein2019]. Ideally, meta-analyses are part of a systematic review. As such, methods are carefully reported upon to ensure the review and data collation are transparent and reproducible [i.e., see the Preferred Reporting Items for Systematic Reviews and Meta-Analyses in Ecology and Evolutionary Biology (PRISMA-EcoEvo) checklist, @ODea2021].

A meta-analysis can have three goals: 1) to provide an overall mean estimate of a treatment effect or relationship, 2) to quantify effect size variance and understand key drivers explaining differences in effects across studies and 3) attempt to identify research gaps and publication biases [@Nakagawa2017; @Gurevitch1999; @Gurevitch2018; @Cooper2009; @Koricheva2013; @Borenstein2019]. Meta-analysing independent studies provides greater statistical power and precision than what any individual study on its own would be able to provide -- particularly given that most empirical studies are already under-powered in many areas of biology [@Forstmeier2017; @Jennions2003; @Button2013]. By expressing study effects on a common scale (i.e., standardised effect size) we can gain broader insight into the direction and efficacy of a particular treatment effect or the strength of a relationship between two variables of interest [@Gurevitch2018; @Koricheva2013]. Meta-analyses have already provided comparative physiologists with powerful insights on pressing global challenges -- from testing whether physiological plasticity can buffer organisms against climate change [e.g, @Seebacher2015] to the degree to which endocrine disrupting chemicals, such as bisphenol A (BPA), impact aquatic organisms [e.g., @WuSeebacher2020].

@@

We included comparative studies / meta-analyses that were focused on physiological questions. Narrative reviews and comments were not considered because they did not collate quantitative data from the literature. Studies needed to have collected physiological trait data, such as hormone levels, thermal physiological parameters (Critical thermal maximum (CTmax), critical thermal minimum (CTmin), thermal preferences, thermal breadth, lethal thermal tolerance limit (LT50), lethal thresholds), metabolic rates (routine, maximum, and minimum metabolic rate, aerobic scope), enzyme reaction rates and concentrations, oxidative stress measures and consequences (e.g., DNA damage, teleomere attrition), performance measurements (e.g., heart rates, sprint speed, swim speed, growth rates), body composition (e.g., fatty acid profiles) and immune traits (e.g., cell function, antibodies). Studies were also included if they measured the relationship between physiological traits and body size (e.g., metabolic scaling), reproductive allocation, or survival across taxa. Studies on behavior were only included if the synthesis was explicitly interested in the proximate physiological drivers of behavioral variation. We also included meta-analyses that quantified plasticity in physiological traits in response to environmental variation. While many studies in comparative physiology collect data as part of a comparative analysis, we focus here only on studies collating data from existing literature because this is the major data source in quantitative syntheses. We excluded quantitative syntheses interested in phenology, bio-mechanics, sexual size dimorphism (i.e., except if the study was interested in sexual differences in physiological traits), those focused on validating physiological methodology, or characterising population genetic structure, gene transcription, and quantitative genetics. There were also meta-analyses that focused on colour evolution which varied in their goals but were most often not focused on physiological questions. As such, these studies were also excluded. We restricted our focus to animal studies because most plant synthesis papers were focused on ecological (e.g., quantified changes in abundance and distribution of species) or agricultural questions (e.g., studies on soil microbial function, nitrogen and phosphorus capture, methane output, and greenhouse gas emissions).
